\chapter{Future Work}
\label{future_work}

\begin{itemize}

\item multiple-level branches are not supported in the current implementation. that is, all collector nodes have the
      root node as their parent and only have leaf nodes as their children.
      \begin{itemize}
      \item support would allow large group configurations to be automatically split into multiple (identically
            configured) branches for improved scalability
      \end{itemize}

\item network failure: \dcamp does not support any fault tolerance for network failures; \dcamp only attempts to recover
from node failures. It is assumed that if (part of) the network goes down, the lack of data from that subnet will
suffice. specifically, \dcamp does not tolerate the split brain syndrome\cite{needed}.

\item time accuracy: The system time among multiple nodes in the system may vary significantly; \dcamp is not meant to
be a high-resolution system with respect to the order of performance data occurrences. It is assumed that NTP provides
sufficient time synchronization across all nodes in the system OR the precise ordering of performance events in the
system is not required.

\item use IOLoop (green events?) instead of polling sockets for messages
      \begin{itemize}
      \item look into using an io loop with callbacks to better manage sockets
      \item possibly, use a single io loop, hosted by the base node service and passed to the roles it starts?
      \item this could greatly increase the scalability of dCAMP and reduce its overhead
      \end{itemize}

\item security --- send salted passphrase with every control message, encrypt all messages, etc.

\end{itemize}

\subsubsection{Metric Extensions}

\textbf{VARIABLE LENGTH DATA}
Does \dcamp need to support arbitrary data lengths?

\textbf{NESTED METRICS}
It could be possible for data messages to contain nested data messages, e.g. average/sum of rates.

\textbf{GROUPINGS}
\dcamp may need a more compact data message format for combining multiple metrics into a single message, e.g. for
aggregation purposes or representing entire branches in the topology.
group by source/type/time?
group start/end frames?

\textbf{HISTOGRAMS}
\dcamp should provide a metric histogram ability. Perhaps this should done at each node or only at the root.

\textbf{METRIC REQUESTS}
The admin can use \dcamp to request metrics on a one-time basis, for example, to enumerate the available disks on each
node. The key here is a metric collection is not based on configuration file but rather on real-time input from the
end-user.

Perhaps there should be some special, e.g. "once", metric collection specifications so config data can be sent up to the
root only at node start.

